Exceedance probability comparison for the simplified Naomi model

### 2023-12-28 src/naomi-simple_exceedance/exceedance.Rmd
Add custom plot of second 90 exceedance to poster #57
Update naming in exceedance notebook (GEB!)

df_second90 <- data.frame(
  prob = c(tmbstan_second90, tmb_second90, aghq_second90),
  method = c(rep("tmbstan", times = n), rep("TMB", times = n), rep("aghq", times = n)),
  index = c(1:n, 1:n, 1:n)
)

@@

plot1 <- ggplot(df_wide_second90, aes(x = tmbstan, y = TMB)) +
  geom_point(shape = 1, alpha = 0.6, col = "#999999") +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "NUTS", y = "TMB") +
  theme_minimal()

plot2 <- ggplot(df_wide_second90, aes(x = tmbstan, y = aghq)) +
  geom_point(shape = 1, alpha = 0.6, col = "#999999") +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "NUTS", y = "PCA-AGHQ") +
  theme_minimal()

@@

mse_change <- round(100 * (mse_aghq_second90 - mse_tmb_second90) / mse_tmb_second90)
rmse_change <- round(100 * (rmse_aghq_second90 - rmse_tmb_second90) / rmse_tmb_second90)
```
There is a change of `r mse_change`% in the MSE moving from `TMB` to `aghq`, and a change of `r rmse_change`% in the RMSE.

@@

```{r}
tmb_fails_second90 <- df_wide_second90 %>%
  mutate(diff = tmbstan - TMB) %>%
  filter(diff > 0.1) %>%
  pull(index)
```
And for `aghq`:

```{r}
aghq_fails_second90 <- df_wide_second90 %>%
  mutate(diff = tmbstan - aghq) %>%
  filter(diff > 0.1) %>%
  pull(index)
```

@@

unique(mf_out_fine[aghq_fails_second90, ]$area_id)
unique(mf_out_fine[aghq_fails_second90, ]$sex)
unique(mf_out_fine[aghq_fails_second90, ]$age_group)
```
Add colour by sex to the above plots:

@@

plot1 <- ggplot(df_wide_second90, aes(x = tmbstan, y = TMB, col = sex)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  scale_color_manual(values = c("#E69F00", "#999999")) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(col = "Sex") +
  guides(col = "none") +
  labs(x = "NUTS", y = "TMB") +
  theme_minimal()

plot2 <- ggplot(df_wide_second90, aes(x = tmbstan, y = aghq, col = sex)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  scale_colour_manual(values = c("#E69F00", "#999999"), labels = c("Female", "Male")) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  labs(x = "NUTS", y = "PCA-AGHQ", col = "Sex") +
  theme_minimal()

ggsave("second90.png", plot1 + plot2, h = 4, w = 6.25)

plot1 + plot2
```
Check the RMSE stratified by sex:

```{r}
df_wide_second90 %>%
  group_by(sex) %>%
  summarise(
    "RMSE TMB (percent)" = 100 * sqrt(mean((tmbstan - TMB)^2)),
    "RMSE aghq (percent)" = 100 * sqrt(mean((tmbstan - aghq)^2))
  )
```

@@

df_1inc <- data.frame(
  prob = c(tmbstan_1inc, tmb_1inc, aghq_1inc),
  method = c(rep("tmbstan", times = n), rep("TMB", times = n), rep("aghq", times = n)),
  index = c(1:n, 1:n, 1:n)
)

@@

plot1 <- ggplot(df_wide_1inc, aes(x = tmbstan, y = TMB)) +
  geom_point(shape = 1, alpha = 0.6, col = "#999999") +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "NUTS", y = "TMB") +
  theme_minimal()

plot2 <- ggplot(df_wide_1inc, aes(x = tmbstan, y = aghq)) +
  geom_point(shape = 1, alpha = 0.6, col = "#999999") +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "NUTS", y = "PCA-AGHQ") +
  theme_minimal()

ggsave("1inc.png", plot1 + plot2, h = 4, w = 6.25)

@@

@@

@@

df_metrics <- df_wide %>%
  select(TMB, aghq, truth = tmbstan, indicator) %>%
  pivot_longer(
    cols = c("TMB", "aghq"),
    names_to = "method",
    values_to = "approximate"
  ) %>%
  group_by(indicator, method) %>%
  summarise(
    rmse = sqrt(mean((truth - approximate)^2)),
    mae = mean(abs(truth - approximate))
  ) %>%
  mutate(method = fct_reorg(method, "TMB" = "TMB", "PCA-AGHQ" = "aghq"))

@@

df_metrics <- df_metrics %>%
  left_join(
    mutate(df_metrics_pct, method = "PCA-AGHQ"),
    by = c("indicator", "method")
  ) %>%
  mutate(
    label = ifelse(
      method == "PCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  ) %>%
  ungroup()

write_csv(df_metrics, "exceedance.csv")

exceedance_manuscript_plot <- df_wide %>%
  pivot_longer(cols = c("TMB", "aghq"), names_to = "method", values_to = "estimate") %>%
  mutate(
    method = fct_reorg(method, "TMB" = "TMB", "PCA-AGHQ" = "aghq"),
    indicator = fct_relevel(indicator, "Second 90", "High incidence")
  ) %>%
  ggplot(aes(x = tmbstan, y = estimate - tmbstan)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.6) +
    scale_colour_manual(values = c("#56B4E9", "#999999"), labels = c("Female", "Male")) +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "Approximate - NUTS", col = "Sex") +
    geom_text(
      data = df_metrics %>% mutate(
        method = fct_relevel(method, "TMB", "PCA-AGHQ"),
        indicator = fct_relevel(indicator, "Second 90", "High incidence")    
      ),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )

@@

exceedance_manuscript_plot_second90 <- df_wide %>%
  filter(indicator == "Second 90") %>%
  pivot_longer(cols = c("TMB", "aghq"), names_to = "method", values_to = "estimate") %>%
  mutate(
    method = fct_reorg(method, "TMB" = "TMB", "PCA-AGHQ" = "aghq")
  ) %>%
  ggplot(aes(x = tmbstan, y = estimate)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.6) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("#56B4E9", "#999999"), labels = c("Female", "Male")) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    geom_text(
    data = df_metrics %>%
      filter(indicator == "Second 90") %>%
      mutate(method = fct_relevel(method, "TMB", "PCA-AGHQ")),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "", col = "Sex") +
    theme_minimal()

ggsave("exceedance-second90.png", exceedance_manuscript_plot_second90, h = 4, w = 6.25)

exceedance_manuscript_plot_1inc <- df_wide %>%
  filter(indicator == "High incidence") %>%
  pivot_longer(cols = c("TMB", "aghq"), names_to = "method", values_to = "estimate") %>%
  mutate(
    method = fct_reorg(method, "TMB" = "TMB", "PCA-AGHQ" = "aghq")
  ) %>%
  ggplot(aes(x = tmbstan, y = estimate)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.6) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("#56B4E9", "#999999"), labels = c("Female", "Male")) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    geom_text(
    data = df_metrics %>%
      filter(indicator == "High incidence") %>%
      mutate(method = fct_relevel(method, "TMB", "PCA-AGHQ")),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "", col = "Sex") +
    theme_minimal()

ggsave("exceedance-1inc.png", exceedance_manuscript_plot_1inc, h = 4, w = 6.25)
```

>>>>

df_second90 <- data.frame(
  prob = c(tmbstan_second90, tmb_second90, aghq_second90),
  method = c(rep("NUTS", times = n), rep("GEB", times = n), rep("GPCA-AGHQ", times = n)),
  index = c(1:n, 1:n, 1:n)
)

@@

plot1 <- ggplot(df_wide_second90, aes(x = NUTS, y = GEB)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "NUTS", y = "GEB") +
  theme_minimal()

plot2 <- ggplot(df_wide_second90, aes(x = NUTS, y = `GPCA-AGHQ`)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(x = "NUTS", y = "GPCA-AGHQ") +
  theme_minimal()

@@

mse_change <- round(100 * (mse_aghq_second90 - mse_tmb_second90) / mse_tmb_second90)
rmse_change <- round(100 * (rmse_aghq_second90 - rmse_tmb_second90) / rmse_tmb_second90)
```

There is a change of `r mse_change`% in the MSE moving from `TMB` to `aghq`, and a change of `r rmse_change`% in the RMSE.

@@

```{r}
tmb_fails_second90 <- df_wide_second90 %>%
  mutate(diff = NUTS - GEB) %>%
  filter(diff > 0.1) %>%
  pull(index)
```

And for `aghq`:

```{r}
aghq_fails_second90 <- df_wide_second90 %>%
  mutate(diff = NUTS - `GPCA-AGHQ`) %>%
  filter(diff > 0.1) %>%
  pull(index)
```

@@

unique(mf_out_fine[aghq_fails_second90, ]$area_id)
unique(mf_out_fine[aghq_fails_second90, ]$sex)
unique(mf_out_fine[aghq_fails_second90, ]$age_group)
```

Add colour by sex to the above plots:

@@

plot1 <- ggplot(df_wide_second90, aes(x = NUTS, y = GEB, col = sex)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  scale_colour_manual(values = c("#0072B2", "black")) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", col = "#E69F00") +
  labs(col = "Sex") +
  guides(col = "none") +
  labs(x = "NUTS", y = "GEB") +
  theme_minimal()

plot2 <- ggplot(df_wide_second90, aes(x = NUTS, y = `GPCA-AGHQ`, col = sex)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("#0072B2", "black"), labels = c("Female", "Male")) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", col = "#E69F00") +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  labs(x = "NUTS", y = "GPCA-AGHQ", col = "Sex") +
  theme_minimal()

ggsave("second90.png", h = 4, w = 6.25)
```

Check the RMSE stratified by sex:

```{r}
df_wide_second90 %>%
  group_by(sex) %>%
  summarise(
    "RMSE GEB (percent)" = 100 * sqrt(mean((NUTS - GEB)^2)),
    "RMSE GPCA-AGHQ (percent)" = 100 * sqrt(mean((NUTS - `GPCA-AGHQ`)^2))
  )
```

@@

df_1inc <- data.frame(
  prob = c(tmbstan_1inc, tmb_1inc, aghq_1inc),
  method = c(rep("NUTS", times = n), rep("GEB", times = n), rep("GPCA-AGHQ", times = n)),
  index = c(1:n, 1:n, 1:n)
)

@@

plot1 <- ggplot(df_wide_1inc, aes(x = NUTS, y = GEB)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", col = "#E69F00") +
  labs(x = "NUTS", y = "GEB") +
  theme_minimal()

plot2 <- ggplot(df_wide_1inc, aes(x = NUTS, y = `GPCA-AGHQ`)) +
  geom_point(shape = 1, alpha = 0.6) +
  coord_fixed(ratio = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", col = "#E69F00") +
  labs(x = "NUTS", y = "GPCA-AGHQ") +
  theme_minimal()

@@

ggsave("1inc.png", h = 4, w = 6.25)

@@

write_csv(df_wide, "exceedance.csv")

@@

df_metrics <- df_wide %>%
  select(GEB, `GPCA-AGHQ`, truth = NUTS, indicator) %>%
  pivot_longer(
    cols = c("GEB", "GPCA-AGHQ"),
    names_to = "method",
    values_to = "approximate"
  ) %>%
  group_by(indicator, method) %>%
  summarise(
    rmse = sqrt(mean((truth - approximate)^2)),
    mae = mean(abs(truth - approximate))
  )

@@

df_metrics <- df_metrics %>%
  left_join(
    mutate(df_metrics_pct, method = "GPCA-AGHQ"),
    by = c("indicator", "method")
  ) %>%
  mutate(
    label = ifelse(
      method == "GPCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  ) %>%
  ungroup()

write_csv(df_metrics, "exceedance-summary.csv")

exceedance_manuscript_plot <- df_wide %>%
  pivot_longer(cols = c("GEB", "GPCA-AGHQ"), names_to = "method", values_to = "estimate") %>%
  mutate(
    indicator = fct_relevel(indicator, "Second 90", "High incidence")
  ) %>%
  ggplot(aes(x = NUTS, y = estimate - NUTS)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.4) +
    scale_colour_manual(values = c("#0072B2", "black"), labels = c("Female", "Male")) +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed", col = "#E69F00") +
    facet_grid(indicator ~ method) +
    labs(x = "Estimate", y = "Absolute difference to NUTS", col = "Sex") +
    geom_text(
      data = df_metrics %>% mutate(
        indicator = fct_relevel(indicator, "Second 90", "High incidence")    
      ),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )

@@

exceedance_manuscript_plot_second90 <- df_wide %>%
  filter(indicator == "Second 90") %>%
  pivot_longer(cols = c("GEB", "GPCA-AGHQ"), names_to = "method", values_to = "estimate") %>%
  ggplot(aes(x = NUTS, y = estimate)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.4) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("#0072B2", "black"), labels = c("Female", "Male")) +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed", col = "#E69F00") +
    geom_text(
    data = df_metrics %>%
        filter(indicator == "Second 90"),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "Absolute difference to NUTS", col = "Sex") +
    theme_minimal()

ggsave("exceedance-second90.png", exceedance_manuscript_plot_second90, h = 3, w = 6.25)

exceedance_manuscript_plot_1inc <- df_wide %>%
  filter(indicator == "High incidence") %>%
  pivot_longer(cols = c("GEB", "GPCA-AGHQ"), names_to = "method", values_to = "estimate") %>%
  ggplot(aes(x = NUTS, y = estimate)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.6) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("#0072B2", "black"), labels = c("Female", "Male")) +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed", col = "#E69F00") +
    geom_text(
      data = df_metrics %>%
        filter(indicator == "High incidence"),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "Absolute difference to NUTS", col = "Sex") +
    theme_minimal()

ggsave("exceedance-1inc.png", exceedance_manuscript_plot_1inc, h = 3, w = 6.25)
```
